### 2019-03-03 report/final-report.tex
results section for simple symmetric synthesis

>>>>

\subsection{Simple Symmetric Synthesis Algorithms}
We implemented simple symmetric lenses as an extension to
Boomerang~\cite{boomerang}.  
We also integrated our synthesis engine into Boomerang, allowing users
to write synthesis tasks 
alongside lens combinators, incorporate synthesis results into manually-written
lenses, and reference previously defined lenses during synthesis. All experiments
were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of 1600 MHz DDR3
running macOS Mojave.

\paragraph*{Benchmark Suite Construction}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted 8 data cleaning benchmarks from Flash Fill~\cite{flashfill}. Flash Fill data
  cleaning tasks are either derived from online help forums or taken from the
  Excel product team. Note that our tool produces bidirectional transformations
  rather than one-way transformers like Flash Fill. We ensure one direction of
  our bidirectional transformers performs the same unidirectional transformation
  as Flash Fill---the other direction is determined from the round-tripping
  laws. None of these benchmarks were bijective.
\item We adapted 29 benchmarks from Augeas~\cite{augeas}. Augeas is a utility
  that bidirectionally converts between Linux configuration files and an
  in-memory dictionaries. In our benchmarks, we synthesize lenses between Linux
  configuration files and serialized versions of Augeas dictionaries. Because
  these benchmarks merely transformed the information into a structured form,
  synthesized lenses were all bijective.
\item We created 11 additional benchmarks 
  derived from real-world examples and/or the bidirectional programming
  literature. These tasks range from synchronizing REST and JSON web resource
  descriptions to synchronizing BibTeX and EndNote citation descriptions. Five
  of these benchmarks were not bijective lenses.
\end{enumerate}

\paragraph*{Synthesizing Correct Simple Symmetric Lenses}

To determine whether the system can synthesize desired lenses, we ran
it interactively on all 48 benchmark tasks, working with the system to
create sufficient examples and provide useful relevance annotations
(\IE, \emph{essential} and \emph{irrelevant} markings, written as
\SRequire and \Skip, respectively).
In all cases, the desired lens was obtained.
%
The majority of the tasks required only a single example and none required
more than three examples to synthesize the desired lens.
%
Providing relevance annotations 
was needed in only 8 of the 48 tasks. In
practice, we found that adding such annotations was quite easy: if manual inspection
of the lens showed there were too few \IdentityLens{}s, and too many
\Disconnect{}s\footnote{
A \Disconnect{} is a lens combinator that marks a place where
information appears in one format but not the 
other: there is a \emph{disconnect} between the two formats. 
}
  or merges, we would add \SRequire annotations. If synthesis took
too long, we would add \Skip annotations.
%
We verified that the default running mode of our synthesis tool
(\SSOpt{}) generated the 
correct lenses the way programmers often validate their programs: we
manually inspected the code and ran unit tests on the synthesized
code. To determine whether the synthesis procedure generated the
correct lens when running in modes other than \SSOpt{}, we compared
generated lens to the lens synthesized by \SSOpt{}.

\paragraph*{Effectiveness of Optimizations in Simple Symmetric Synthesis}

Having determined appropriate examples and annotations for the 48
benchmarks, we evaluate the performance of the system by measuring the running
time of our algorithm in two modes:
\vskip .5ex
\begin{tabulary}{\linewidth}{rL}
  \SSOpt{}: & Run the symmetric synthesis algorithm with all optimizations enabled.\\
  \SSNCOpt{}: & Run the symmetric synthesis algorithm with no compositional synthesis enabled.\\
\end{tabulary}\\
\noindent
Compositional synthesis allows users to break a benchmark into a
series of smaller synthesis tasks, whose solutions are utilized in more complex
synthesis procedures. Compositional synthesis (\SSOpt{} mode)
allows our system to scale to arbitrarily large and complex formats; measuring
it shows the responsiveness of the system when used as
intended. \SSNCOpt{} mode, which synthesizes a complete lens all at once,
provides a useful experimental stress test for the system.

\begin{figure}
  \centering
  \includegraphics{times}
  \vspace{-2ex}
  \caption{Number of
    benchmarks that can be solved by a given algorithm in a 
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
    existing lenses. The symmetric synthesis algorithm is able to complete all
    benchmarks in under 30 seconds elapsed total time. Without compositional
    synthesis it is able to complete 26. Each benchmark specification includes
    source and target (potentially annotated) regular expressions, and between
    one and three sufficient examples.}
  \label{fig:times}
\end{figure}

For each benchmark in the suite and each mode, we ran the system with a timeout
of 60 seconds, averaging the result over 5 runs. Figure~\ref{fig:times}
summarizes the results of these tests. We find that our algorithm is able to
synthesize all of the benchmarks in under 30 seconds. Without compositional
synthesis, the synthesis algorithm is able to solve 26 out of 48 problem instances.

% Section~\ref{subsec:optimizations} that users may tune the termination metric.
% Success on this set of benchmarks requires we tune that termination condition
% (e.g. for some benchmarks, allowing the system flexibility to spend more time
% considering more of the search space). In

\paragraph*{Slowdown Compared to Bijective Synthesis}

To compare to the existing bijective synthesis algorithm, we run our symmetric
synthesis algorithm on the original Optician benchmarks, comprised of
39 bijective synthesis tasks.\footnote{We had to slightly alter four of these
benchmarks, either by providing additional examples or by adding \SRequire annotations.
Without these alterations, symmetric synthesis yielded a lens that fit
the specification but that was undesired.}
To perform this comparison, we synthesized lenses in two modes:
\vskip 2ex
\begin{tabulary}{\linewidth}{rL}
  \BSOpt{}: & The existing bijective synthesis
              algorithm with all optimizations enabled.\\
  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
\end{tabulary}\\[2ex]
\noindent
For each benchmark, we ran it in both modes with a
timeout of 60 seconds and averaged the result over 5 runs.
Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
average, \SSOpt{} took 1.3 times (0.5 seconds) longer to complete than
\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
depth 3 into a serialized version of the structured dictionary representation
used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
complete.

\begin{figure}[t]
  \includegraphics{times_bijective}
  \vspace{-2ex}
  \caption{Number of benchmarks that can be solved by a given algorithm in a
    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
    \BSOpt{} is the full bijective lens synthesis algorithm.}
  \label{fig:times_bijective}
\end{figure}

\paragraph*{The Effects of Heuristics and Relevance Annotations}
We evaluate the usefulness of (1) our information-theoretic metric, (2) our
termination heuristic and (3) our
relevance annotations.  To this end, we run our program in several different modes:\\[2ex]
\begin{tabulary}{\linewidth}{rL}
  \AnyOpt{}: & Ignore the information-theoretic metric (\IE, all valid
               lenses have cost 0). \\ 
  \FLOpt{}: &  Return the first highest ranked lens \GreedySynth returns (\IE,
              ignore the termination heuristic). \\
  \CCOpt{}: & Replace our information-theoretic cost metric with one where
              the cost of the lens is the number of disconnects plus the number
              of merges.\\
  \NSOpt{}: & Ignore all \Skip annotations in the SRE specifications. \\
  \NROpt{}: & Ignore all \SRequire annotations in the SRE specifications. \\[2ex]
\end{tabulary}

\noindent
We experimented with the \CCOpt{} mode to determine whether the
complexity of the information-theoretic measure was really
needed. Related work on string transformations has often used simpler
measures such as ``avoid constants'' that align with, but are simpler
than our measures. The \CCOpt{} mode is an example of such a simple
measure---it operates by counting disconnects, which put a complete
stop to information transfer, and merges, which
eliminate the information in a union.

\begin{figure}[t]
  \includegraphics{metrics_importance}
  \vspace{-2ex}
  \caption{Number of benchmarks that synthesize the correct lens by a given
    algorithm. \AnyOpt{} provides no notion of cost, and merely returns the
    first lens it finds that satisfies the specification. \FLOpt{} provides a
    notion of cost to \GreedySynth, but once a satisfying lens is greedily
    found, that lens is returned. \CCOpt{} synthesizes lenses, where the cost of
    a lens is the number of disconnects plus the number of merges. \NSOpt{}
    ignores all \Skip annotations while running the algorithm. \NROpt{} ignores
    all \SRequire annotations while running the algorithm.}
  \label{fig:metric}
\end{figure}

Figure~\ref{fig:metric} summarizes the result of these experiments. The data
reveal that the information-theoretic metric is critical for finding the correct
lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode. The
termination condition is also quite important. When running in \FLOpt{} mode,
the algorithm only discovers 5 lenses, which shows that the first class that
contains a satisfying lens is rarely the correct class. However, our algorithm
is not perfect and fails when either it is very difficult to find the desired
lens (necessitating \SRequire) or when a large amount of data is projected
(necessitating \Skip). Without any annotations, our algorithm finds the correct
lens for 40 of our 48 benchmarks; eight required relevance annotations to find
the correct lens.
% end evaluation
